---
title: "Rabies 1"
author: "Kính"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme:
      bootswatch: minty
---

Some settings

```{r setup, include=FALSE}
here::i_am("code.Rmd")
source('~/Documents/libs.r')
```

Mexico

```{r, main_map, include=F}
mex <- ne_countries(scale = 110, type = "map_units", returnclass = "sf", country = "Mexico")
mexico <- ne_states(country = "Mexico", returnclass = "sf")
main_states <- char(Campeche, Chiapas, "Quintana Roo", "Yucatán")
states <- filter(mexico, name_id %in% main_states)
```

## CENARPRECE { .tabset}

```{r readxls, echo=F, include=F}
# Data raw
bovine <- readxl::read_excel(here("CENAPRECE/Data/Aggregated_clean/bovine_rabies_2018-20.xlsx")) |>
    mutate(across(where(is.character), str_to_lower))
# DT::datatable(bovine, caption = 'Bovine')
animal <- readxl::read_excel(here("CENAPRECE/Data/Aggregated_clean/animal_agressions_2018-2020.xlsx")) |>
    mutate(across(where(is.character), str_to_lower)) |> 
    rename(ent = state, med = medical_unit)
# DT::datatable(animal, caption = 'Animal')
pep <- readxl::read_excel(here("CENAPRECE/Data/Aggregated_clean/pep_2018-2020.xlsx")) |>
    mutate(across(where(is.character), str_to_lower)) |>
    rename(ent = state, med = medical_unit)
# DT::datatable(pep, caption = 'PEP')
h_units <- readxl::read_excel(here("HEALTH_UNITS/ESTABLECIMIENTO_SALUD_202212.xlsx")) |>
    select(
        ent = `NOMBRE DE LA ENTIDAD`,
        ent_key = `CLAVE DE LA ENTIDAD`,
        mun = `NOMBRE DEL MUNICIPIO`,
        mun_key = `CLAVE DEL MUNICIPIO`,
        loc = `NOMBRE DE LA LOCALIDAD`,
        loc_key = `CLAVE DE LA LOCALIDAD`,
        med = `NOMBRE DE LA UNIDAD`,
        lat = `LATITUD`, lon = `LONGITUD`
    ) |>
    filter(stringr::str_to_lower(ent) %in% str_to_lower(main_states)) |>
    mutate(across(everything(), str_to_lower))
# DT::datatable(h_units[1L:500, ], caption = 'Health units')
```

### Missing location

Merge and filter those without lat and lon, then search with `osm`.

```{r, echo = FALSE}
missing_ll <- animal |>
    left_join(h_units, c('ent', 'med'), multiple = 'all')|>
    filter(is.na(lat)) |>
    count(ent, med, mun, loc)
missing_ll <- pep |>
    left_join(h_units, c('ent', 'med'), multiple = 'all') |>
    filter(is.na(lat)) |>
    count(ent, med, mun, loc)
    bind_rows(missing_ll) |>
    distinct()
```

```{r, arg_gis, eval = FALSE, echo = F}
# This run before compile output as it's a long running process.
dir.create('prep')
library(tidygeocoder)
try_ll <- missing_ll %>%
    mutate(
        medx = if_else(is.na(loc), med, paste(med, ', ', loc)),
        medx = if_else(is.na(mun), medx, paste(medx, ', ', mun)),
        medx = paste0(medx, ', ', ent, ', ', 'mexico'),
        medx = stringr::str_squish(medx)
    ) |> 
    geocode(
        address = medx,
        method = "arcgis",
        full_results = TRUE
    )
saveRDS(try_ll, here('prep/geocoded.rds'))
```

### Fixed location {.active}

```{r, map_animal, echo = F}
try_ll <- readRDS(here("prep/geocoded.rds"))

animal %>%
    left_join(h_units, "medical_unit") %>%
    mutate(across(c(lat, lon), as.numeric)) %>%
    left_join(try_ll %>% select(lat2 = lat, long, medical_unit, arcgis_address), "medical_unit") %>%
    mutate(
        lat = if_else(is.na(lat) & !is.na(lat2), lat2, lat),
        lon = if_else(is.na(lon) & !is.na(long), long, lon),
        fixed = if_else(!is.na(lat2), T, F)
    ) %>%
    filter(!is.na(lat)) %>%
    st_as_sf(coords = c("lon", "lat"), crs = sf::st_crs(4326)) %>%
    allot(animalZ)

ggplot() +
    geom_sf(data = mexico) +
    geom_sf(aes(color = factor(fixed), size = cases, alpha = cases), animalZ) +
    labs(color = "Fixed missing")

ggplot() +
    geom_sf(data = states) +
    geom_sf(
        aes(color = factor(fixed), size = cases, alpha = cases), 
        data = st_crop(animalZ, states)
    ) +
    geom_sf_text(aes(label = name_id), data = states) +
    labs(title = "Main states", color = "Fixed missing")
```

### Todo - check fixed location

For the moment, ignore those missing. @Ivonne can you check

- if the identified ones are correct (there is address obtained)
- if the nonidentified ones can be fixed/located (e.g. QRSSA0…)

```{r, print_missing, results = 'asis'}
DT::datatable(try_ll)
```

---

## SINAVE pdfs { .tabset}

```{r, eval = FALSE, echo=F}
# Something like this can be done, but area is not easy to define/different between files.
library(tabulizer)
tabel <- extract_tables(
    here(
        "SINAVE/Datos-20210914T123636Z-001/Datos/Boletin epi-SINAVE/2010_sem52.pdf"), 
    pages = c(24), 
    guess = FALSE,
    method = "stream",
    area = list(c(252.9551, 114.4653, 637.6408, 495.1102))
)
# Use `tabula` directly is faster. Then read in here
```

Mordedura read

```{r, echo = F}
mordedura <- list.files('~/Documents/WIP/Papers/rabies/data/SINAVE/mordedura/', full.names = T) %>%
    lapply(read.csv) %>%
    data.table::rbindlist(idcol = 'src', use.names = FALSE)

cname <- char(
    src, loc,
    perr_cur_sem, perr_cur_m, perr_cur_f, perr_pre_acum,
    mami_cur_sem, mami_cur_m, mami_cur_f, mami_pre_acum,
    serp_cur_sem, serp_cur_m, serp_cur_f, serp_pre_acum
)

years <- list.files("~/Documents/WIP/Papers/rabies/data/SINAVE/mordedura/") %>%
    basename() %>%
    gsub(".*(20[0-9]{2}).*", "\\1", .)

mordedura %>%
    mutate(src = years[src]) %>%
    rename_with(~cname) %>%
    pivot_longer(-c(src, loc), names_sep = '_', names_to = char(animal, time, type)) %>%
    mutate(
        count = gsub('\\s', '', value),
        count = if_else(count == '-', "0", count),
        count = as.numeric(count), 
        src  = as.numeric(src),
        year = case_when(
            time == 'cur' ~ src,
            time == 'pre' ~ src - 1,
        )
    ) %>%
    allot(mordedura)
```

> @Ivonne please check randomly some number if the PDF reading is correct.

### Data

```{r, echo = F, results = 'asis'}
DT::datatable(mordedura)
```

### Time trend

> Seems something is wrong with 2022 report?

```{r, echo = F}
dir.create('fig', F)
mordedura %>%
    filter(time == "cur", type %in% char(m, f)) %>%
    group_by(animal, type, year) %>%
    summarise(n = sum(count)) %>%
    ggplot() +
    geom_line(aes(year, n, color = type)) +
    facet_wrap(~animal, scales = "free_y") +
    labs(title = 'Bite count time trend', caption = 'from SINAVE pdfs')
```

```{r, echo = F}
dir.create('fig', F)
mordedura %>%
    filter(time == "cur", type %in% char(m, f)) %>%
    group_by(loc, animal, type, year) %>%
    summarise(n = sum(count)) %>%
    ggplot() +
    geom_line(aes(year, n, color = factor(loc))) +
    facet_grid(vars(animal), vars(type), scales = "free_y") +
    labs(title = "Bite count time trend", caption = "from SINAVE pdfs") +
    guides(color = "none") +
    ggrepel::geom_text_repel(
        aes(year, n, label = loc),
        data = ~ filter(., year == 2023), 
        seed = 78
    )
```

### Map

```{r, echo=F}
mordedura %>%
    rename(name_id = loc) %>%
    left_join(mexico %>% select(name_id), "name_id") %>%
    allot(mordeduraZ)

mordeduraZ %>%
    filter(
        time == "cur", type == "m", 
        year %in% c(2009, 2015, 2020, 2023)
    ) %>%
    ggplot() +
        geom_sf(aes(fill = count, geometry = geometry), color = 'grey95') +
        facet_grid(vars(animal), vars(year)) +
        labs(title = "Mordedura - M") +
        scale_fill_viridis_c()
```

## SIRVENA {.tabset}

First animal file does not have colnames.

### A data
```{r, echo = F}
A <- list.files(here("SIRVERA/SIREVA/"), "SIRVERA_A", recursive = 1, full.names = T) %>%
    mapply(readxl::read_excel, ., col_names = c(F, T, T), SIMPLIFY = F)
colnames(A[[1]]) <- colnames(A[[2]])
A <- bind_rows(A)
A %<>%
    select(
        country = `País de Ocurrencia`,
        adm1 = `ADM1 de Ocurrencia`,
        adm2 = `ADM2 de Ocurrencia`,
        date = `Fecha de la Notificación`,
        year = `Año`,
        mon = Mes, 
        week = Semana,
        variant = `Variante Antigénica`,
        tested = `Realizo teste de laboratorio?`,
        animal = `Especie`,
        animal_group = `Grupo de Especie`
    ) %>%
    filter(country == "México")
DT::datatable(A, caption = "Animal")
```
over time
```{r, echo=F}
A %>%
    filter(animal %in% c("Perro", "Bovino")) %>%
    group_by(adm1, animal, year) %>%
    count() %>%
    ggplot(aes(year, n, color = adm1)) +
    geom_line() +
    facet_wrap(~animal, scales = "free_y") +
    guides(color = "none") +
    labs(title = "A data - colors by states") +
    ggrepel::geom_text_repel(aes(label = adm1), ~ filter(., year == 2018), seed = 78)
```

### H data

```{r, echo = F}
H <- list.files(here("SIRVERA/SIREVA/"), "SIRVERA_H", recursive = 1, full.names = T) %>%
    lapply(readxl::read_excel) %>%
    bind_rows() %>%
    select(
        country = `País de Ocurrencia`,
        adm1 = `ADM1 de Ocurrencia`,
        adm2 = `ADM2 de Ocurrencia`,
        date_not = `Fecha de Notificación`,
        date_exp = `Fecha de Exposición`,
        date_die = `Fecha del Óbito`,
        year = `Año`,
        mon = `Mes`,
        week = `Semana`,
        age = Edad,
        country_exp = `País de Exposición`,
        adm1_exp = `ADM1 de Exposición`,
        adm2_exp = `ADM2 de Exposición`,
        year_exp = `Año de Exposición`,
        mon_exp = `Mes de Exposición`,
        wek_exp = `Semana de Exposición`,
        variant = `Variante Antigénica`,
        tested = `Realizo teste de laboratorio?`,
        animal = `Especie Transmisora`,
        animal_group = `Grupo de Especie Transmisora`
    ) %>%
    filter(country == "México")
DT::datatable(H, caption = 'H data')
```

over time

```{r, echo=F}
H$age %>% hist(breaks = 30)

H %>%
    group_by(adm1, year, animal) %>%
    count() %>%
    ggplot(aes(year, n, color = adm1)) +
    geom_line() +
    facet_wrap(~animal, scales = "free_y") +
    guides(color = "none") +
    labs(title = "H data - colors by states") +
    ggrepel::geom_text_repel(aes(label = adm1), ~ filter(., year == 2018), seed = 78)
```