---
title: "Rabies 1"
author: "Kính"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    theme:
      bootswatch: minty
---

Some settings

```{r setup, include=FALSE}
here::i_am("code.Rmd")
source('~/Documents/libs.r')
```

Mexico

```{r, main_map, include=F}
mex <- ne_countries(scale = 110, type = "map_units", returnclass = "sf", country = "Mexico")
mexico <- ne_states(country = "Mexico", returnclass = "sf")
main_states <- char(Campeche, Chiapas, "Quintana Roo", "Yucatán")
states <- filter(mexico, name_id %in% main_states)
# https://www.geoboundaries.org/index.html#getdata
adm2 <- st_read("adm2/geoBoundaries-MEX-ADM2_simplified.shp")
```

## CENARPRECE { .tabset}

```{r readxls, echo=F, include=F}
# Data raw
bovine <- readxl::read_excel(here("CENAPRECE/Data/Aggregated_clean/bovine_rabies_2018-20.xlsx")) |>
    mutate(across(where(is.character), str_to_lower))
# DT::datatable(bovine, caption = 'Bovine')
animal <- readxl::read_excel(here("CENAPRECE/Data/Aggregated_clean/animal_agressions_2018-2020.xlsx")) |>
    mutate(across(where(is.character), str_to_lower)) |> 
    rename(ent = state, med = medical_unit)
# DT::datatable(animal, caption = 'Animal')
pep <- readxl::read_excel(here("CENAPRECE/Data/Aggregated_clean/pep_2018-2020.xlsx")) |>
    mutate(across(where(is.character), str_to_lower)) |>
    rename(ent = state, med = medical_unit)
# DT::datatable(pep, caption = 'PEP')
h_units <- readxl::read_excel(here("HEALTH_UNITS/ESTABLECIMIENTO_SALUD_202212.xlsx")) |>
    select(
        ent = `NOMBRE DE LA ENTIDAD`,
        ent_key = `CLAVE DE LA ENTIDAD`,
        mun = `NOMBRE DEL MUNICIPIO`,
        mun_key = `CLAVE DEL MUNICIPIO`,
        loc = `NOMBRE DE LA LOCALIDAD`,
        loc_key = `CLAVE DE LA LOCALIDAD`,
        med = `NOMBRE DE LA UNIDAD`,
        lat = `LATITUD`, lon = `LONGITUD`
    ) |>
    filter(stringr::str_to_lower(ent) %in% str_to_lower(main_states)) |>
    mutate(across(everything(), str_to_lower))
# DT::datatable(h_units[1L:500, ], caption = 'Health units')
```

### Missing location

Merge and filter those without lat and lon, then search with `osm`.

```{r, echo = FALSE}
missing_ll <- animal |>
    left_join(h_units, c('ent', 'med'), multiple = 'all')|>
    filter(is.na(lat)) |>
    count(ent, med, mun, loc)
missing_ll <- pep |>
    left_join(h_units, c('ent', 'med'), multiple = 'all') |>
    filter(is.na(lat)) |>
    count(ent, med, mun, loc)
    bind_rows(missing_ll) |>
    distinct()
```

```{r, arg_gis, eval = FALSE, echo = F}
# This run before compile output as it's a long running process.
dir.create('prep')
library(tidygeocoder)
try_ll <- missing_ll %>%
    mutate(
        medx = if_else(is.na(loc), med, paste(med, ', ', loc)),
        medx = if_else(is.na(mun), medx, paste(medx, ', ', mun)),
        medx = paste0(medx, ', ', ent, ', ', 'mexico'),
        medx = stringr::str_squish(medx)
    ) |> 
    geocode(
        address = medx,
        method = "arcgis",
        full_results = TRUE
    )
saveRDS(try_ll, here('prep/geocoded.rds'))
```

### Fixed location {.active}

ENT and MED are matched but not LOC

- just picked one of them now, obs. can be a wrong LOC!
- some is not in the three states

```{r, map_animal, echo = F}
try_ll <- readRDS(here("prep/geocoded.rds"))

anpep <- bind_rows(
    left_join(animal, h_units, c("ent", "med"), multiple = "first"),
    left_join(pep, h_units, c("ent", "med"), multiple = "first")
) |> 
    mutate(across(c(lat, lon), as.numeric)) |>
    left_join(
        select(try_ll, ent, med, mun, loc, lat2 = lat, long, arcgis_address),
        c("ent", "med", "mun", "loc")
    ) |> 
    mutate(
        lat = if_else(is.na(lat) & !is.na(lat2), lat2, lat),
        lon = if_else(is.na(lon) & !is.na(long), long, lon),
        fixed = if_else(!is.na(lat2), T, F)
    ) |>
    filter(!is.na(lat))

anpep |>
    st_as_sf(coords = c("lon", "lat"), crs = sf::st_crs(4326)) |>
    ggplot() +
    geom_sf(aes(color = type, size = cases)) +
    labs(title = 'with err')

anpep |>
    filter(lon > -95) |>
    st_as_sf(coords = c("lon", "lat"), crs = sf::st_crs(4326)) |>
    ggplot() +
    geom_sf(aes(color = type, size = cases)) +
    labs(title = "remove err")

anpep %<>% filter(lon > -95) 
```

## SINAVE pdfs { .tabset}

Mordedura read

```{r mordedura_read, echo = F}
mordedura <- list.files('~/Documents/WIP/Papers/rabies/data/SINAVE/mordedura/', full.names = T) |>
    lapply(read.csv) |>
    data.table::rbindlist(idcol = 'src', use.names = FALSE)

cname <- char(
    src, loc,
    perr_cur_sem, perr_cur_m, perr_cur_f, perr_pre_acum,
    mami_cur_sem, mami_cur_m, mami_cur_f, mami_pre_acum,
    serp_cur_sem, serp_cur_m, serp_cur_f, serp_pre_acum
)

years <- list.files("~/Documents/WIP/Papers/rabies/data/SINAVE/mordedura/") |>
    basename() %>% 
    gsub(".*(20[0-9]{2}).*", "\\1", .)

mordedura  <-  mordedura |>
    mutate(src = years[src]) |>
    rename_with(~cname) |>
    pivot_longer(-c(src, loc), names_sep = '_', names_to = char(animal, time, type)) |>
    mutate(
        count = gsub('\\s', '', value),
        count = if_else(count == '-', "0", count),
        count = as.numeric(count), 
        src  = as.numeric(src),
        year = case_when(
            time == 'cur' ~ src,
            time == 'pre' ~ src - 1,
        )
    )
```

> Seems something is wrong with 2022 report?

### Map

```{r, echo=F}
mordeduraZ <- mordedura |>
    rename(name_id = loc) |>
    left_join(mexico |> select(name_id), "name_id") %>%
    filter(str_to_lower(name_id) %in% str_to_lower(main_states))

nn <- mordeduraZ |>
    filter(time == "cur", type %in% c("m", "f"), animal != "serp") %>%
    group_by(name_id, animal) %>%
    summarise(n = sum(count), n_year = length(unique(year)), n = n / n_year)

mordeduraZ %>% count(year)
mmm <- states %>%
    left_join(select(nn, name = name_id, animal, n)) %>%
    filter(animal == "mami") %>%
    ggplot() +
    geom_sf(aes(fill = n)) +
    ggrepel::geom_text_repel(
        aes(label = name, geometry = geometry),
        size = 3, color = "white",
        stat = "sf_coordinates",
        bg.color = "black",
        bg.r = 0.15,
        # min.segment.length = 0
    ) +
    scale_fill_gradientn(colours = hcl.colors(4, 'Heat 2', rev = T)) +
    theme(panel.grid = element_blank(), axis.text = element_blank()) +
    labs(
        x = "", y = "",
        fill = "Número medio anual\nde mordeduras", title = "otros mamíferos"
    )
prm <- states %>%
    left_join(select(nn, name = name_id, animal, n)) %>%
    filter(animal == "perr") %>%
    ggplot() +
    geom_sf(aes(fill = n)) +
    ggrepel::geom_text_repel(
        aes(label = name, geometry = geometry), 
        size = 3, color = 'white', 
        stat = "sf_coordinates",
        bg.color = "black", 
        bg.r = 0.15,
        # min.segment.length = 0
        ) +
    scale_fill_viridis_c() +
    theme(panel.grid = element_blank(), axis.text = element_blank()) +
    labs(
        x = "", y = "",
        fill = "Número medio anual\nde mordeduras", 
        title = "Perros"
    )
mmm+prm
# ggsave('./fig/sinave.pdf', width = 9, height = 3.5)
```

## INEGI

```{r inegi_var}
# remotes::install_github("wmgeolab/rgeoboundaries")
# library(rgeoboundaries)
lcld <- list.files('/Users/knguyen/Seafile/INEGI/', 'LOCALIDAD', recursive = 1, full.names = 1)
lcld <- readr::read_csv(lcld, col_select = c(
    ENT, MUN, LOC, TIPOLOC, ACT_PRIN, TAMLOC
))

itdt <- list.files('/Users/knguyen/Seafile/INEGI/', 'ITER_', recursive = 1, full.names = 1)
itdt <- readr::read_csv(itdt, col_select = c(
    ENT = ENTIDAD, NOM_ENT, MUN, NOM_MUN, LOC, NOM_LOC,
    lon = LONGITUD, lat = LATITUD, pop = POBTOT,
    VPH_C_SERV, VPH_NDEAED
)) |>
    filter(str_detect(NOM_MUN, "Total", T)) |>
    filter(str_detect(NOM_LOC, "Total", T))

inegi <- itdt |>
    full_join(lcld, c("ENT", "MUN", "LOC")) |>
    filter(!is.na(lon), !is.na(lat)) |>
    mutate(
        lon = as.numeric(sp::char2dms(lon, "°")),
        lat = as.numeric(sp::char2dms(lat, "°")),
        TIPOLOC = if_else(TIPOLOC %in% c('9', '99'), NA_character_, TIPOLOC)
    )
```

There are `r inegi |> filter(is.na(lon) | is.na(lat)) |> nrow()` obs. with missing lon or lat.

```{r label_inegi}
act_prin_lb <- c(
    `1` = "Grow/harvest agricultural",
    `2` = "Animal husbandry",
    `3` = "Exploitation of mines",
    `4` = "Cutting/planting trees",
    `5` = "Gathering wild herbs/plants",
    `6` = "Fishing/hunting animals",
    `7` = "Making handicrafts",
    `8` = "Others",
    `9` = "Not specified",
    `NA` = "No information"
)

tipoloc_lb <- c(
    `01` = "Ranch/farm",
    `02` = "Hamlet",
    `03` = "Traditional illage",
    `04` = "Industrialised housing",
    `05` = "Self-produced housing",
    `06` = "Worker's hostel/camp",
    `07` = "Commercial/service establishments",
    `08` = "Industrial zone/park",
    `09` = "Other",
    `99` = "Not specified"
)

act_prin_lb <- c(
    `1` = "Cultivo o cosecha de productos agrícolas",
    `2` = "Crianza de animales",
    `3` = "Explotación de minas",
    `4` = "Corte o siembra de árboles",
    `5` = "Recolección de hierbas o plantas silvestres",
    `6` = "Pesca o caza de animales",
    `7` = "Elaboración de artesanías",
    `8` = "Otra actividad económica",
    `9` = "No especificado",
    `NA` = "Sin información"
)

tipoloc_lb <- c(
    `01` = "Rancho o finca",
    `02` = "Caserío",
    `03` = "Poblado tradicional",
    `04` = "Desarrollo habitacional de vivienda industrializada",
    `05` = "Barrio, colonia o fraccionamiento de vivienda autoproducida",
    `06` = "Albergue o campamento de trabajadores",
    `07` = "Establecimientos comerciales y de servicios",
    `08` = "Zona o parque industrial",
    `09` = "Otro tipo",
    `99` = "No especificado"
)
```


### h3 - correlation on hexagon grid of 36km2

```{r h3_aggregate_dem}
res <- 6
inegi$h3 <- geo_to_h3(inegi[, c("lat", "lon")], res)

# find most frequent type in a hexagon
# if missing get what ever values of the region
tipo <- inegi |>
    group_by(h3, TIPOLOC) |>
    count() |>
    # fill NA within h3 group
    ungroup() |>
    group_by(h3) |>
    fill(TIPOLOC, .direction = "downup") |>
    # now get max freq within group
    ungroup() |>
    group_by(h3) |>
    slice_max(n, with_ties = FALSE) |>
    select(h3, TIPOLOC)

# if missing get what ever values of the region
actprin <- inegi |>
    group_by(h3, ACT_PRIN) |>
    count() |>
    # fill NA within h3 group
    ungroup() |>
    group_by(h3) |>
    fill(ACT_PRIN, .direction = "downup") |>
    # now get max freq within group
    ungroup() |>
    group_by(h3) |>
    slice_max(n, with_ties = FALSE) |>
    select(h3, ACT_PRIN)

# up to 2 hexes/loc -> NA in the aggregated
inegi |>
    count(ENT, MUN, LOC, h3) %>%
    count(n)

pop <- inegi |>
    group_by(h3) |>
    # group_by(ENT, MUN, LOC, h3) |>
    summarise(pop = sum(pop, na.rm = T))

# so let it be, if we have the matched hex, the rate is correct, else we will fill
# those after joining with cases data

econ <- inegi |>
    mutate(econ = as.numeric(VPH_C_SERV)) |>
    group_by(h3) |>
    summarise(econ = sum(econ, na.rm = T))

statesh3 <- states |> fill_polygon_h3(res)
stateshex <- statesh3 |> h3_to_geo_boundary_sf() |> rename(h3 = h3_index)

# get rate
fd <- stateshex |>
    left_join(tipo) |>
    left_join(actprin) |>
    left_join(pop) |>
    left_join(econ)

fd %<>% mutate(ACT_PRIN = factor(ACT_PRIN))
```

```{r plot_tm, fig.dim = c(7, 7)}
tm_shape(fd) +
    tm_polygons(c("pop", "econ", "TIPOLOC", "ACT_PRIN")) +
    tm_facets(sync = TRUE, ncol = 2) +
    tm_style("cobalt")
```

Only animal and pep are on fine-scale

```{r case_corr}
anpep$h3 <- geo_to_h3(anpep[, c("lat", "lon")], res)

# count cases
anpeph3 <- anpep |>
    group_by(h3, year, type) |>
    summarise(cases = sum(cases, na.rm = T))

fdc <- fd %>%
    left_join(anpeph3, multiple = "all")

fdc %>%
    filter(!is.na(type)) %>%
    ggplot(aes(cases)) +
    facet_wrap(~type) +
    geom_histogram()

fdc %<>% mutate(case = if_else(is.na(cases), 0, cases), lcase = log1p(case))

rawmap <- fdc %>%
    ggplot() +
    geom_sf(data = states, color = "grey90", fill = '#0c263a') +
    geom_sf(aes(fill = case), color = "transparent", linewidth = 0) +
    scale_fill_gradientn(colors = c('transparent', viridisLite::magma(30, begin = 0.3, direction = 1)), trans = 'log1p') +
    geom_sf_text(
        aes(label = name, geometry = geometry), 
        data = states,
        size = 3, 
        stat = "sf_coordinates", 
        nudge_x = c(1.4, -1.2, 1, 1),
        nudge_y = c(0, 1, -1, 1.2),
    ) +
    theme(
        legend.key.height = unit(2, 'cm'),
        legend.position = c(0, .5),
        panel.grid = element_blank(), axis.text = element_blank()) +
    labs(
        x = "", y = "",
        fill = "Número de casos\n(2020)"
    )
# ggsave('./fig/raw_map.pdf', rawmap, width = 7.5, height = 7)
```

## Some boxes 

```{r boxcorr}
options(
    ggplot2.discrete.colour = thematic::okabe_ito()
    ggplot2.discrete.fill = thematic::okabe_ito()
)

cols <- gen_colors(okabe, 12)

f1 <- fdc %>%
    filter(!is.na(type), !is.na(TIPOLOC)) %>%
    filter(type != "pep") %>%
    # mutate(act = if_else(act == 9, '8', act)) %>%
    ggplot() +
    geom_boxplot(aes(TIPOLOC, cases, fill = TIPOLOC), alpha = .48) +
    geom_jitter(aes(TIPOLOC, cases, fill = TIPOLOC), position = position_jitter(.1)) +
    scale_y_continuous(trans = "log1p") +
    scale_x_discrete(labels = tipoloc_lb) +
    scale_fill_manual(values = cols[1:5]) +
    coord_flip() +
    labs(y = "Incidences (log1p)", x = "Location type") +
    guides(fill = "none") +
    theme(strip.background = element_rect("grey95", "grey99"))

f2 <- fdc %>%
    filter(!is.na(type), !is.na(ACT_PRIN)) %>%
    filter(type != "pep") %>%
    mutate(ACT_PRIN = if_else(ACT_PRIN == '9', '8', ACT_PRIN)) %>% 
    ggplot() +
    geom_boxplot(aes(ACT_PRIN, cases, fill = ACT_PRIN), alpha = .48) +
    geom_jitter(aes(ACT_PRIN, cases, fill = ACT_PRIN), position = position_jitter(.1)) +
    scale_y_continuous(trans = "log1p") +
    scale_x_discrete(labels = act_prin_lb) +
    scale_fill_manual(values = cols[6:12]) +
    coord_flip() +
    labs(y = 'Incidences (log1p)', x = "Activity type") +
    guides(fill = 'none') +
    theme(strip.background = element_rect('grey95', 'grey99'))
f1/f2 + plot_layout(height = c(.42, .58))

# ggsave('./fig/inc_act_tipo.pdf', width = 9, height = 6)
```


## Zero-inflated model

```{r smooth}
library(INLA)
idt <- filter(fdc, type != "pep", !is.na(pop)) %>% mutate(idx = 1:n())
nrow(idt)

result = inla(
    cases ~ 1 + TIPOLOC + ACT_PRIN + factor(year) + f(idx),
    family = "zeroinflatedpoisson1",
    control.predictor = list(link = 1),
    control.compute = list(config = TRUE),
    data = idt
)

summary(result)

# result$misc$configs$contents 
sp <- INLA::inla.posterior.sample(1000, result)

fun = function(...) {
    c(
        i0 = exp(Intercept),
        TIPOLOC02 = (1 - theta[1]) * exp(Intercept + TIPOLOC02),
        TIPOLOC03 = (1 - theta[1]) * exp(Intercept + TIPOLOC03),
        TIPOLOC05 = (1 - theta[1]) * exp(Intercept + TIPOLOC05),
        TIPOLOC07 = (1 - theta[1]) * exp(Intercept + TIPOLOC07),
        ACT_PRIN2 = (1 - theta[1]) * exp(Intercept + ACT_PRIN2),
        ACT_PRIN3 = (1 - theta[1]) * exp(Intercept + ACT_PRIN3),
        ACT_PRIN4 = (1 - theta[1]) * exp(Intercept + ACT_PRIN4),
        ACT_PRIN5 = (1 - theta[1]) * exp(Intercept + ACT_PRIN5),
        ACT_PRIN6 = (1 - theta[1]) * exp(Intercept + ACT_PRIN6),
        ACT_PRIN7 = (1 - theta[1]) * exp(Intercept + ACT_PRIN7),
        ACT_PRIN8 = (1 - theta[1]) * exp(Intercept + ACT_PRIN8),
        ACT_PRIN9 = (1 - theta[1]) * exp(Intercept + ACT_PRIN9),
        y2019 = exp(Intercept + `factor(year)2019`),
        y2202 = exp(Intercept + `factor(year)2020`)
    )
}
f2 = inla.posterior.sample.eval(fun, sp)

f2 %>%
    apply(1, \(x) quantile(x, probs = c(.5, .025, .975))) %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    pivot_longer(-rowname) %>%
    pivot_wider(names_from = rowname, values_from = value) %>%
    rename(rowname = name) %>% 
    rownames_to_column %>%
        mutate(rowname = case_when(
            rowname == "TIPOLOC02" ~ tipoloc_lb[2],
            rowname == "TIPOLOC03" ~ tipoloc_lb[3],
            rowname == "TIPOLOC05" ~ tipoloc_lb[5],
            rowname == "TIPOLOC07" ~ tipoloc_lb[7],
            rowname == "ACT_PRIN2" ~ act_prin_lb[2],
            rowname == "ACT_PRIN3" ~ act_prin_lb[3],
            rowname == "ACT_PRIN4" ~ act_prin_lb[4],
            rowname == "ACT_PRIN5" ~ act_prin_lb[5],
            rowname == "ACT_PRIN6" ~ act_prin_lb[6],
            rowname == "ACT_PRIN7" ~ act_prin_lb[7],
            rowname == "ACT_PRIN8" ~ act_prin_lb[8],
            rowname == "ACT_PRIN9" ~ act_prin_lb[9],
            rowname == "factor(year)2019" ~ "2019 vs 2018",
            rowname == "factor(year)2020" ~ "2020 vs 2018",
            rowname == "zero-probability parameter for zero-inflated poisson_1" ~ "zero",
            rowname == "Precision for idx" ~ "prec"
        )) %>%
        filter(rowname != "zero") %>%
        filter(rowname != "prec") %>%
        filter(!is.na(rowname)) %>%
        mutate(type = c(
            rep('Tipo de actividad (ref: agricultura)', 4), 
            rep('Tipo de localidad (ref: rancho o finca)', 7), 
            rep('Año', 2))) %>% 
        allot(tmpm)

p1 <- 
tmpm %>%
    ggplot(aes(rowname, `0.5quant`, color = type)) +
    geom_hline(yintercept = 0, linetype = 'dotted', color = 'grey50') +
    geom_linerange(aes(ymin = `0.025quant`, ymax = `0.975quant`)) +
    geom_point(shape = 21, fill = "white", size = 2) +
    facet_wrap(~type, 3, scales = 'free_y', ) +
    labs(
        x = '', y = 'Estimaciones de parámetros\ne intervalo de confianza de 95%',
        # title = "Zeroinflated Negative Binomial Type I",
        # caption = 'Complete-case regression without spatial correlation'
    ) +
    theme(
        legend.position = "none", 
        axis.title.x = element_text(size = 9),
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()
    ) +
    coord_flip()
load_all("~/Code/R/ktools")
lo <- p1 %>% facet_unwrap(plot = F)
lo[[2]]$labels$y <- lo[[1]]$labels$y <- ""
lo[[2]] / lo[[1]] / lo[[3]] + plot_layout(heights = c(.6, .4, .2), guides = 'collect')
ggsave('./fig/simple_model.pdf', width = 7, height = 4.5)
```

OK, if get all these filled for all hexagons maybe we can predict a little.

### Test k-ring smooth of covariates

```{r kring_smooth_tipo}
tipo <- kring_smooth(fdc, "h3", "TIPOLOC", 3)
fdc$tipo <- tipo

lm <- scale_color_manual(
    values = thematic::okabe_ito(10),
    limits = unique(fdc$TIPOLOC)
)
bf <- ggplot(fdc, aes(fill = TIPOLOC)) +
    geom_sf() +
    lm +
    labs(title = "Before k-ring smoothing")
at <- ggplot(fdc, aes(fill = tipo)) +
    geom_sf() +
    lm +
    labs(title = "After k-ring smoothing")
bf + at
```

```{r kring_smooth_act}
fdc %<>% mutate(ACT_PRIN = as.character(ACT_PRIN))
actprin <- kring_smooth(fdc, "h3", "ACT_PRIN", 4)
fdc$actprin <-actprin 

lm <- scale_color_manual(
    values = thematic::okabe_ito(10),
    limits = unique(fdc$actprin )
)
bf <- ggplot(fdc, aes(fill =actprin )) +
    geom_sf() +
    lm +
    labs(title = "Before k-ring smoothing")
at <- ggplot(fdc, aes(fill =actprin )) +
    geom_sf() +
    lm +
    labs(title = "After k-ring smoothing")
bf + at
```

```{r inla_smooth, cache=TRUE, eval = FALSE}
idt <- fdc %>%
    filter(type != "pep") %>%
    # drop_na(any_of(char(tipo, actprin, idd))) %>% 
    mutate(idd = as_numeric(h3), idx = 1:n())
nrow(idt)

H <- spdep::poly2nb(idt, queen = T)
spdep::nb2INLA("hex.adj", H)
graph <- inla.read.graph("hex.adj")

fml <- cases ~ 1 +
    tipo + actprin + 
    # factor(year) + 
    f(idx) +
    f(idd,
        model = "besag", 
        constr = T, scale.model = T, 
        graph = graph
    )

result1 = inla(fml,
    family = "zeroinflatedpoisson1",
    data = idt,
    control.inla = list(int.strategy = 'eb'),
    control.predictor = list(link = 1)
)

# idt$fitted <- result1$summary.fitted.values$`0.5quant`
idt$fitted <- result1$summary.fitted.values$mean

idt %>%
    ggplot(aes(cases, fitted)) +
    geom_point()

idt %>%
    ggplot(aes(fill = fitted)) +
    geom_sf() +
    geom_sf(data = states, fill = "transparent", color = 'white', inherit.aes = F) +
    scale_fill_viridis_c(trans = "log") +
    labs(
        title = "Smoothed estimated of all region",
        subtitle = "Use spatial correlation but not covariates"
    ) +
    theme(panel.grid = element_blank())
idt %>%
    ggplot(aes(fill = cases)) +
    geom_sf() +
    geom_sf(data = states, fill = "transparent", color = 'white', inherit.aes = F) +
    scale_fill_viridis_c(trans = "log") +
    labs(
        title = "Original data",
        subtitle = "Use spatial correlation but not covariates"
    ) +
    theme(panel.grid = element_blank())
```
